---
title: "ReproducibleResearch_PeerAssignment"
author: "Grant McClure"
date: "2/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in the data

```{r}
library("data.table")
library(ggplot2)
setwd("/Users/grmcclure/OneDrive - Red Ventures/Documents/Projects/dsa/")
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = paste0(getwd(), '/repdata%2Fdata%2Factivity.zip'), method = "curl")
unzip("repdata%2Fdata%2Factivity.zip",exdir = "data")

data <- read.csv("data/activity.csv")  
```

## What is mean total number of steps taken per day?
1) Calculate the total number of steps taken per day
2) If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
3) Calculate and report the mean and median of the total number of steps taken per day

``` {r}
#total mean steps per day
stepsPerDay <- aggregate(steps ~ date, data, sum)
hist(stepsPerDay$steps, main = paste("Total Steps Each Day"),col = "blue",xlab="Number of Steps")

meanOfSteps <- mean(stepsPerDay$steps)
meanOfSteps
medianOfSteps <- median(stepsPerDay$steps)
medianOfSteps
```

## What is the average daily activity pattern?

1) Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

``` {r}
#average daily activity pattern

stepsPerInterval <- aggregate(steps ~ interval, data, mean)
plot(stepsPerInterval$interval,stepsPerInterval$steps, type="l", xlab="Interval", 
     ylab="Number of Steps",main="Average Number of Steps by day for each Interval")

maxInterval <- stepsPerInterval[which.max(stepsPerInterval$steps),1]
maxInterval
```

## Imputing missing values

```{r}
#imputing missing values

#number of missing values
nrow(data[is.na(data$steps),])

#get rid of na's by replacing with mean
StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
fixSteps <- numeric()
for (i in 1:nrow(data)) {
  row <- data[i, ]
  if (is.na(row$steps)) {
    steps <- subset(StepsAverage, interval == row$interval)$steps
  } else {
    steps <- row$steps
  }
  fixSteps <- c(fixSteps, steps)
}

finalData <- data
finalData$steps <- fixSteps

stepsAllIn <- aggregate(steps ~ date, data = finalData, sum, na.rm = TRUE)
hist(stepsAllIn$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
#Create Histogram to show difference. 
hist(stepsPerDay$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)

#calculate mean
meanAllIn <- mean(stepsAllIn$steps)
meanAllIn
medianAllIn <- median(stepsAllIn$steps)
medianAllIn

medianDiff <- medianAllIn - medianOfSteps
medianDiff
meanDiff <- meanAllIn - meanOfSteps
meanDiff

#difference seen in the 10000--15000 bucket, which is what we'd expect since we replaced nulls with the mean
```

## Are there differences in activity patterns between weekdays and weekends?
``` {r}
#differences between weekends and weekdays?
finalData$weekday <- c("weekday")
finalData[weekdays(as.Date(finalData[, 2])) %in% c("Saturday", "Sunday", "samedi", "dimanche", "saturday", "sunday", "Samedi", "Dimanche"), ][4] <- c("weekend")
table(finalData$weekday == "weekend")
finalData$weekday <- factor(finalData$weekday)

finalData_weekend <- subset(finalData, finalData$weekday == "weekend")
finalData_weekday <- subset(finalData, finalData$weekday == "weekday")

mean_finalData_weekday <- tapply(finalData_weekday$steps, finalData_weekday$interval, mean)
mean_finalData_weekend <- tapply(finalData_weekend$steps, finalData_weekend$interval, mean)

library(lattice)
df_weekday <- NULL
df_weekend <- NULL
df_final <- NULL
df_weekday <- data.frame(interval = unique(finalData_weekday$interval), 
                         avg = as.numeric(mean_finalData_weekday), 
                         day = rep("weekday", length(mean_finalData_weekday)))
df_weekend <- data.frame(interval = unique(finalData_weekend$interval), 
                         avg = as.numeric(mean_finalData_weekend), 
                         day = rep("weekend", length(mean_finalData_weekend)))
df_final <- rbind(df_weekday, df_weekend)

xyplot(avg ~ interval | day, data = df_final, layout = c(1, 2), 
       type = "l", ylab = "Number of steps")
```

